Applying Natural Language Processing and Machine Learning techniques to aid in the diagnosis of Mild Cognitive Impairment and Early Dementia - A Systematic Review

### 2019-09-16 Source/Paper 2 - Systematic Review/Systematic Review.tex
updated SR

\keywords{MCI \and Alzheimer's Disease \and Machine Learning \and Diagnosis} 
\bigskip
\begin{abstract}
This is the paper's abstract \ldots
\end{abstract}

\section{Introduction}
Dementia has been identified as one of those fast growing difficulties facing the world. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
\par
A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}. As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages \cite{Prince2015} and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD. At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}. I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
\par
There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}. However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}. Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression. Another area of focus for research of this nature is the process of collecting appropriate language samples. Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task. Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD. There is some evidence to support this. Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control. 
\par
The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease. The potential impact of this research in this area is immense. Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease however the absence of a single test and the complexity of AD can create significant delays in diagnosis. Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
\par 
The remainder of this article is organized as follows. Section~\ref{methodology} gives an account of the process of this Systematic Review. Our results are described in Section~\ref{results}. We discuss the results and implications in Section~\ref{discussion}. Finally, Section~\ref{conclusions} gives the conclusions.

@@

	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & Authors \\ \hline
	2 & Abnormalities of Connected Speech in Semantic Dementia vs Alzheimer’s diseases & Authors \\ \hline
	3 & Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech & Authors \\ \hline
	4 & Alignment of spoken narratives for automated neuropsychological assessment & Authors \\ \hline
	5 & Automatic detection and rating of dementia of alzheimer type through lexical analysis of spontaneous speech & Authors \\ \hline
	6 & Changes in Style in Authors with Alzheimer’s Disease & Authors \\ \hline
	7 & Computer-based evaluation of Alzheimer's disease and mild cognitive impairment patients during a picture description task & Authors \\ \hline
	8 & Computerized assessment of syntactic complexity in Alzheimer's disease: A case study of Iris Murdoch's writing & Authors \\ \hline
	9 & Computerized neuropsychological assessment in mild cognitive impairment based on natural language processing-oriented feature extraction & Authors \\ \hline
	10 & Contrastive conversational analysis of language production by Alzheimer's and control people& Authors \\ \hline
	11 & Declines in Connected Language Are Associated with Very Early Mild Cognitive Impairment: Results from the Wisconsin Registry for Alzheimer’s Prevention & Authors \\ \hline
	12 & Deep-Deep Neural Network Language Models for Predicting Mild Cognitive Impairment & Authors \\ \hline
	13 & Domain Adaptation for Detecting Mild Cognitive Impairment & Authors \\ \hline
	14 & Diagnosing people with dementia using automatic Conversation Analysis & Authors \\ \hline
	15 & Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia & Authors \\ \hline
	16 & Enriching Complex Networks withWord Embeddings for Detecting Mild Cognitive Impairment from Speech Transcripts & Authors \\ \hline
	17 & Features and Machine Learning Classification of Connected Speech Samples from Patients with Autopsy Proven Alzheimer’s Disease with and without Additional Vascular Pathology & Authors \\ \hline
	18 & Formulaic Language in People with Probable Alzheimer’s Disease: A Frequency-Based Approach & Authors \\ \hline
	19 & Language Analysis of Speakers with Dementia of the Alzheimer’s Type & Authors \\ \hline
	20 & Language analytics for assessing brain health: Cognitive impairment, depression and pre-symptomatic Alzheimer's disease & Authors \\ \hline
	21 & Learning Linguistic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams & Authors \\ \hline
	22 & Learning Predictive Linguistic Features for Alzheimer’s Disease and related Dementias using Verbal Utterances & Authors \\ \hline
	23 & Linguistic Features Identify Alzheimer's Disease in Narrative Speech & Authors \\ \hline
	24 & Longitudinal detection of dementia through lexical and syntactic changes in writing: a case study of three British novelists & Authors \\ \hline
	25 & Microlinguistic aspects of the oral narrative in patients with Alzheimer’s disease & Authors \\ \hline
	26 & Multilingual word embeddings for the assessment of narrative speech in mild cognitive impairment & Authors \\ \hline
	27 & Natural Language Features for Detection of Alzheimer's Disease in Conversational Speech & Authors \\ \hline
	28 & NLP-Oriented Contrastive Study of Linguistic Productions of Alzheimer's and Control People & Authors \\ \hline
	29 & Predicting mild cognitive impairment from spontaneous spoken utterances & Authors \\ \hline
	30 & Predicting probable Alzheimer's disease using linguistic deficits and biomarkers & Authors \\ \hline
	31 & Propositional Idea Density in women’s written language over the lifespan: Computerized analysis & Authors \\ \hline
	32 & Sentence segmentation in narrative transcripts from neuropsychological tests using recurrent convolutional neural networks & Authors \\ \hline
	33 & Speech processing approach for diagnosing dementia in an early stage & Authors \\ \hline
	34 & Spoken language biomarkers for detecting cognitive impairment & Authors \\ \hline
	35 & Talk2Me: Automated linguistic data collection for personal assessment & Authors \\ \hline
	36 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & Authors \\ \hline
	37 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & Authors \\ \hline
	39 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & Authors \\ \hline
	40 & Vector-space topic models for detecting Alzheimer's disease & Authors \\ \hline
	41 & Vocabulary Size in Speech May Be an Early Indicator of Cognitive Impairment. & Authors \\ \hline

>>>>

\keywords{MCI \and Alzheimer's Disease \and Machine Learning \and Diagnosis} 
\bigskip
\begin{abstract}
Dementia is a growing problem which faces the world today with the total worldwide cost of dementia last year projected to be in the region of one trillion US dollars. Given the huge impact Dementia has, a lot of research has been focused on finding ways to improve the early diagnosis of dementia and Mild Cognitive Impairment. Equally, our ability to process language and find patterns of decline using the techniques of Natural Language Processing and Machine Learning has vastly increased over the recent times. This represents an opportunity to harness these techniques to find patterns in cognitive decline that could indicate the presence of the early stages of dementia.  This paper looks at body of research focused on the analysis of cognitive decline in those suspected to have dementia. Specifically,   \ldots
\end{abstract}

\section{Introduction}
Dementia has been identified as one of the fast growing difficulties facing the world. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
\par
A lot of work has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cognitive Impairment (MCI) with research focused on two areas - identifying biological markers and analyzing the cognitive decline of those who are suspected to have the disease \cite{Taler2008}. As described above \cite{Prince2015}, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilize technology wherever possible to aid clinicians in the detection of MCI and AD. At the present time diagnosis is typically conducted at memory clinics by trained clinicians \cite{Boschi2017}. I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
\par
There is a large body of research that looks at language deterioration in those suspected to have MCI or AD \cite{Taler2008, Boschi2017}. However there is conflicting evidence in these studies about which declining language factors are associated with MCI and AD \cite{Taler2008, Boschi2017}. Research therefore, should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression. Another area of focus for research of this nature is the process of collecting appropriate language samples. Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task. Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD. There is some evidence to support this. Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control. 
\par
The question to be addressed in this systematic review is how has the field of machine learning and natural language processing addresses language deterioration in the diagnosis of Mild Cognitive Impairment and Early Alzheimer's Disease. The potential impact of this research in this area is immense. Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease however the absence of a single test and the complexity of AD can create significant delays in diagnosis. Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
\par 
The remainder of this article is organized as follows. Section~\ref{methodology} gives an account of the process of this Systematic Review. Our results are described in Section~\ref{results}. We discuss the results and implications in Section~\ref{discussion}. Finally, Section~\ref{conclusions} gives the conclusions.

@@

	1 & A Comparison of Syntax, Semantics, and Pragmatics in Spoken Language among Residents with Alzheimer's Disease in Managed-Care Facilities & C. Guinn, B. Singer, A. Habash \\ \hline
	2 & Aided Diagnosis of Dementia Type through Computer-Based Analysis of Spontaneous Speech & W. Jarrold, B. Peintner, D. Wilkins, D. Vergryi, C. Richey, M.L. Gorno-Tempini, J. Ogar \\ \hline
	3 & Automatic detection and rating of dementia of alzheimer type through lexical analysis of spontaneous speech & C. Thomas, V. Keselj, N. Cercone, K. Rockwood, E. Asp \\ \hline
	4 & Changes in Style in Authors with Alzheimer’s Disease & G. Hurst, V.W. Feng \\ \hline
	5 & Computer-based evaluation of Alzheimer's disease and mild cognitive impairment patients during a picture description task & L. Hernandez-Dominguez, S. Ratte, G. Sierra-Martinez, A. Roche-Bergua \\ \hline
	6 & Computerized neuropsychological assessment in mild cognitive impairment based on natural language processing-oriented feature extraction & C. Fang, P. Janwattanapong, H. Martin, M. Cabrerizo, A. Barreto, D. Loewenstein, R. Duara, M. Adjouadi \\ \hline
	7 & Deep-Deep Neural Network Language Models for Predicting Mild Cognitive Impairment & S.O. Orimaye, J. S-M. Wong, J. S. G. Fernandez \\ \hline
	8 & Deep language space neural network for classifying mild cognitive impairment and Alzheimer-type dementia & S.O. Orimaye, J. S-M. Wong, C.P. Wong \\ \hline 
	9 & Diagnosing people with dementia using automatic Conversation Analysis & B. Mirheidari, D. Blackburn, M. Reuber, T. Walker, H. Christensen \\ \hline
	10 & Enriching Complex Networks withWord Embeddings for Detecting Mild Cognitive Impairment from Speech Transcripts & L. B. dos Santos, E. A. Correa Jr, O. N. Oliveira Jr, D.R. Amancio, L. L. Mansur, S. M. Alusio \\ \hline
	11 & Features and Machine Learning Classification of Connected Speech Samples from Patients with Autopsy Proven Alzheimer’s Disease with and without Additional Vascular Pathology & V. Rentoumi, L. Raoufian, S. Ahmed, C. A. de Jager, P. Garrard \\ \hline
	12 & Language Analysis of Speakers with Dementia of the Alzheimer’s Type & C. Guinn, A. Habash \\ \hline
	13 & Learning Linguistic Biomarkers for Predicting Mild Cognitive Impairment using Compound Skip-grams & S. O. Orimaye, J. S-M. Wong, K. Y. Tai, C. P. Wong \\ \hline
	14 & Learning Predictive Linguistic Features for Alzheimer’s Disease and related Dementias using Verbal Utterances & S.O. Orimaye, J. S-M. Wong, K. J. Golden \\ \hline
	15 & Linguistic Features Identify Alzheimer's Disease in Narrative Speech & K. C. Fraser, J. A. Meltzer, F. Rudzicz \\ \hline
	16 & Longitudinal detection of dementia through lexical and syntactic changes in writing: a case study of three British novelists & X. Le, I. Lancashire, G. Hirst, R. Jokel \\ \hline
	17 & Multilingual word embeddings for the assessment of narrative speech in mild cognitive impairment &  K. C. Fraser, K. L. Fors, D. Kokkinakis \\ \hline
	18 & Predicting mild cognitive impairment from spontaneous spoken utterances & M. Asgari, J. Kaye, H. Dodge \\ \hline
	19 & Predicting probable Alzheimer's disease using linguistic deficits and biomarkers & S. O. Orimaye, J. S-M. Wong, K. J. Golden, C. P. Wong, I. N. Soyiri \\ \hline
	20 & Sentence segmentation in narrative transcripts from neuropsychological tests using recurrent convolutional neural networks & M. V. Treviso, C. Shulby, S. M. Alusio \\ \hline
	21 & Social Markers of Mild Cognitive Impairment: Proportion of Word Counts in Free Conversational Speech & H. H. Dodge, N. Mattek, M. Gregor, M. Bowman, A. Seelye, O. Ybarra, M. Asgari, J. A. Kaye \\ \hline
	22 & Spoken language biomarkers for detecting cognitive impairment & T. Alhanai, R. Au, J. Glass \\ \hline
	23 & Spoken Language Derivied Measures for Detecting Mild Cognitive Impairment & B. Roark, M. Mitchell, J-P. Hosom, K. Hollingshead, J. A. Kaye \\ \hline
	24 & Talk2Me: Automated linguistic data collection for personal assessment & M. Komeili, C. Pou-Prom, D. Liaqat, K. C. Fraser, M. Yancheva, F. Rudzicz \\ \hline
	25 & The Effect of Heterogeneous Data for Alzheimer’s Disease Detection from Speech & A. Balagopalan, J. Novikova, F. Rudzicz, M. Ghassemi \\ \hline
	26 & Toward the Automation of Diagnostic Conversation Analysis in Patients with Memory Complaintss & B. Mirheidari, D. Blackburn, K. Harkness, T. Walker, A. Venneri, M. Reuber, H. Cristensen \\ \hline
	27 & Tracking Discourse Complexity Preceding Alzheimer's Disease Diagnosis: A Case Study Comparing the Press Conferences of Presidents Ronald Reagan and George Herbert Walker Bush & V. Berisha, S. Wang, A. LaCross, J. Liss \\ \hline
	28 & Vector-space topic models for detecting Alzheimer's disease & M. Yancheva, F. Rudzicz \\ \hline
